---
title: "Assignment 4"
author: "Gracie White"
date: "November 12, 2018"
always_allow_html: yes
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# Install necessary packages
library(tidyverse)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(extrafont)
library(ggrepel)
library(vcdExtra)
library(effsize)
library(leaflet)
library(car)


# Load Datasets
lobster_size_df <- read_csv("lobster_size_abundance.csv")
traps_df <- read_csv("lobster_traps.csv")

```
##INTRODUCTION
#####
  The spiny lobster, *Panulirus interruptus*, is an integral part of the the California kelp forests. A keystone species, the lobsters exert top down control on urchins, and accordingly, the macroalgae that the urchins consume^3^. While research is still being done on the ecosystem dynamics of this predator/prey relationship in the Santa Barbara area, other locations have documented trophic cascades detrimental to kelp forests stemming from excess fishing pressure on the Spiny Lobster^3^. 
  
  In addition to ecological importance, the Spiny Lobster also supports a multi-million dollar commercial fishery, and millions more in recreational consumer spending, the third most valuable fishery in area (behind the market squid and Dungeness Crab fisheries)^1,2^. For example, the total 2016 ex-vessel value of the California lobster fishery came to $13,691,364, and brought in an additional estimated $33-$40 million in consumer spending from recreational fishing, diving, and eco-tourism^1^. 

	The State of California has been in charge of managing this fishery for over a century, and while recent regulations seem to have stabilized populations, many are concerned that the heavy fishing of the last century has caused an overall decrease in both size and abundance^3^. If so, then the value of the commercial fishery, and the health of the kelp forests, may be at risk^3^.

	The 2016-2017 season had the lowest catch of the last ten fishing seasons, despite an increase in abundance predicted with the Pacific Decadal Oscillation’s warmer waters^2^. The drop in catch was thus unexpected by the Department of Fish and Wildlife. Another 2016-2017 anomaly noticed by researchers was a shift in geographic location. More catch than ever before originated from the Channel Islands, as opposed to the rest of southern California^2^. These anomalies underscore the importance of a robust fisheries management plan, with continued analysis of its effectiveness. 
	
Current management is centered around a limited-entry approach. The fishery is only open from October to March, meant to prevent any harvesting during the spawning season^1^. Recreational fishers are also required to have a specific lobstering license in order to participate, and only a certain number of commercial licenses are awarded each season^2^. To further protect lobsters of spawning age, California also enacted an 82.6mm minimum size limit for carapace length. In theory, this limitation should allow lobsters to reproduce for a year or two before reaching the size limit^1^. 

Additionally, In 2012, the State of California set up 50 new marine protected areas (MPAs) under the Marine Life Protection Act (MLPA) in the the southern part of the state^1^. These MPAs were implemented specifically to conserve fishery resources, such as providing “safe zones” for the Spiny Lobster to reproduce without any fishing pressure^1^. This was not without controversy, as some argued that the creation of MPAs would only serve to increase fishing pressure elsewhere, in non-MPAs, to the detriment of those ecosystems^1^. 

The following report analyzes data between at five Long-Term Ecological Research (LTER) Sites in the Santa Barbara Area, Arroyo Quemado (AQUE), Naples Reef (NAPL), Mohawk Reef (MOHK), Isla Vista (IVEE), Carpinteria (CARP)^4^. Two of the sites, Isla Vista and Naples Reef, were both established as MPAs in 2012. This report will look at the trends in lobster size, overall abundance, and fishing pressure between MPAs and non-MPAs between 2012 and 2017, and suggest potential next steps for monitoring the California Spiny Lobster Fishery. 

##DATA AND METHODS

#####
  Data was provided by Santa Barbara Coastal Long Term Ecological Research Project, coordinated by Dan Reed^4^. Starting in August of 2012, divers recorded the number and size of lobsters in four 300m^2^ transects at each SBC LTER site. For the collection of data on fishing pressure, observers with the SBC LTER project counted the number of trap floats in defined areas of each study site. Each float corresponds to one baited lobster trap under the surface. As the Naples Reef and Isla Vista sites are designated MPAs, there were no observed floats at either site for the duration of the study. Data was recorded every two-four weeks during the fishing season^4^. Figure 1 shows the geographic location of each SBC LTER sampling site considered in this report. 
  
 	The data collected was compiled into two tables, one on fishing pressure, and one on lobster size and abundance. Population statistics were analyzed for significant differences in lobster size, abundance, and fishing pressure between MPA and non-MPA sites. Lobster size by site in 2017 was compared with an omnibus ANOVA (alpha = 0.05 unless otherwise indicated), followed by a Tukey’s HSD post hoc test. Changes in lobster size between 2012 and 2017 at each site was tested with a student’s t-test, after an F-Test showed significant evidence to suggest that samples had equal variances. Finally, the proportion of lobsters above and below the minimum size limit was compared across all sites using a chi-square test, and looking at the standardized residuals. All statistical analysis and graphics were performed in R Statistical Software (V 1.1.456). 


##RESULTS AND DISCUSSION

*Lobster Abundance and Fishing Pressure*
```{r, echo=FALSE}
# Fishing Pressure
# First, filter the data so we only keep relevant info
traps5_df <- traps_df %>%
  select(YEAR, MONTH, FISHING_SEASON, SITE, TRAPS) %>%
  filter(SITE == "CARP" | SITE == "AQUE" | SITE == "MOHK") %>%
  group_by(SITE, YEAR) 
  
# Summary table of lobster traps each year by site  
trap5_sumtable <- summarize(traps5_df, med_traps = median(TRAPS), max_traps = max(TRAPS), mean_traps = mean(TRAPS), sd_traps = sd(TRAPS), sample_size = length(TRAPS), sum_traps = sum(TRAPS))


# Line graph of mean lobster traps by year for each site
traps_line_g <- ggplot(trap5_sumtable, aes(x = YEAR, y = mean_traps, group = SITE, color = SITE)) +
  geom_line(size = 1, alpha = 0.8) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  labs(x = "Year", y = "Observed Traps", title = "Average Annual Observed Lobster Traps at 3 Non-MPA LTER Sites (2012-2017)") + scale_colour_brewer(palette = "YlOrRd")

traps_line_g

# looks like fishing pressure is decreasing over time. The drop after 2016 is probably a result of the decreased abundance seen in the next figure? Note that there are no traps in the MPA sites. 


sum_traps_graph <- ggplot(trap5_sumtable, aes(x = YEAR, y = sum_traps, group = SITE)) + 
  geom_col(position = "dodge", aes(fill = SITE)) + scale_fill_brewer(palette = "YlOrRd") + theme_classic() + scale_y_continuous(expand = c(0,0)) + scale_x_continuous(expand = c(0,0)) + labs(x = "Year", y = "Observed Traps", title = "Total Annual Observed Lobster Traps at 3 Non-MPA LTER Sites (2012-2017)")
sum_traps_graph
  
```

```{r, echo=FALSE}

# Filter sites 
lobcount_allyears_df <- lobster_size_df %>% 
  filter(SITE == "AQUE"| SITE == "NAPL"| SITE == "MOHK"| SITE == "IVEE"| SITE == "CARP", COUNT != "0") %>% 
  select(YEAR, SITE, COUNT) %>% 
  group_by(SITE, YEAR)
#add up all lobster observations by site and year
lobcount <- aggregate(COUNT~SITE+YEAR, data = lobcount_allyears_df, sum) 
#Create column to specify if site is an MPA or not  
lobcount <- mutate(lobcount, MPA = ifelse(SITE == "IVEE" | SITE == "NAPL", "MPA", "No MPA")) %>% 
  group_by(SITE, YEAR)

lobcount

# Line Graph of Mean abundance by site
#traps_line_g <- ggplot(lobcount, aes(x = YEAR, y = mean_traps, group = SITE, color = SITE)) +
  #geom_line(size = 1, alpha = 0.8) +
  #theme_classic() +
  #scale_y_continuous(expand = c(0,0)) +
  #scale_x_continuous(expand = c(0,0)) +
 # labs(x = "Year", y = "Observed Traps", title = "Average Annual Observed Lobster Traps at 3 Non-MPA LTER Sites (2012-2017)") + scale_colour_brewer(palette = "YlOrRd")


# Stacked column graph of total abundance by site
abundance_line_gcol <- ggplot(lobcount, aes(x = YEAR, y = COUNT, group = SITE)) +
  geom_col(position = "dodge", aes(fill = SITE)) +
  labs(x = "Year", y = "Number of Lobsters Observed", title = "Spiny Lobster Abundance in the Santa Barbara Channel (2012-2017)") +
  scale_y_continuous(expand = c(0,0), breaks = seq(0, 900, 100),limits=c(0,750)) +
  scale_x_continuous(expand = c(0,0)) +
  theme_classic() + scale_fill_brewer(palette = "YlOrRd") +
  facet_grid(~ MPA, space = "fixed")


abundance_line_gcol


#Abundance has recently increased in IVEE and CARP, but decreased or stayed similar at other sites. Why is there a peak in 2015, and an overall decrease in 2016? Was there an event that we should be aware of? Why is CARP increasing so much when it's not an MPA, and conversely, why is the NAPL MPA decreasing? Perhaps an El Niño event? A change in fishery regulations? A lobster disease? 
```


*Mean Lobster Size in 2017*
```{r, echo=FALSE}
# Reformat size data into tidy format
# Filter sites and omit counts of "0"
lsize_sites_df <- lobster_size_df %>% 
  filter(SITE == "AQUE"| SITE == "NAPL"| SITE == "MOHK"| SITE == "IVEE"| SITE == "CARP", COUNT != "0", YEAR == "2017") %>% 
  select(YEAR, SITE, SIZE, COUNT)


# Expand Rows
lsize_expanded <- uncount(lsize_sites_df, weights = COUNT, .remove = FALSE)

```


```{r, echo=FALSE}
# One variable (carapace length (mm)), 5 factors
# Perform one-sided ANOVA

leveneTest(SIZE ~ SITE, data = lsize_expanded)

# Identify hypotheses
# H0: Means across all groups are equal
# HA: At least two group means differ significantly
size_aov <- aov(SIZE ~ SITE, data = lsize_expanded)
summary(size_aov)

# Sample means are this different because they were most likely drawn for different population
# H0: Means across all groups are equal
# HA: At least two group means differ significantly

# p-value is <0.05 -> Reject the null.  There is a significant difference in means.  
# Sample means are this different because they were most likely drawn for different population
# Use post-hoc testing to see which sites differ

# Identify hypotheses
# H0:  There is no significant difference in pairwise means
# HA: There is a significant difference in pairwise means
size_pht <- TukeyHSD(size_aov)
size_pht

# Extract labels and factor levels from Tukey post-hoc 
# p-value is <0.05 for sites (NAPL-CARP), (NAPL-IVEE)
# Reject the null for sites (NAPL-CARP), (NAPL-IVEE)

# Create a box-plot for visual mean comparison
size_bp <- ggplot(lsize_expanded, aes(x = SITE, y = SIZE)) +
  geom_boxplot(aes(fill = SITE), color = "black") +
  ggtitle("Lobster Carapace Lengths at 5 LTER Sites in the Santa Barbara Channel, 2017") +
  xlab("Site") +
  ylab("Carapace Length (mm)") +
  theme_classic() +
  scale_fill_brewer(palette = "YlOrRd")
size_bp

sumtablexls <- read_csv("meansdssumtable.csv")
# Create a data summary table to compare means
size_sum_df <- lsize_expanded %>% 
  group_by(SITE) %>% 
  summarize(count = n(), 
    mean = mean(SIZE),
    sd = sd(SIZE))
size_sum_df

sum_size_table <- kable(size_sum_df)
```

*Changes in Lobster Size from 2012-2017 in MPA and non-MPA sites*
```{r, echo=FALSE}
# MPA Sites: IVEE, NAPL
# Non-MPA Sites: AQUE, CARP, MOHK

# Reformat size data into tidy format
# Filter sites and omit counts of "0"
lsize_sitesmpa_df <- lobster_size_df %>% 
  filter(SITE == "AQUE"| SITE == "NAPL"| SITE == "MOHK"| SITE == "IVEE"| SITE == "CARP", COUNT != "0",
         YEAR == "2012"| YEAR == "2017") %>% 
  select(YEAR, SITE, SIZE, COUNT)


# Expand Rows
lsizempa_expanded <- uncount(lsize_sitesmpa_df, weights = COUNT, .remove = FALSE)

# For exploration purposes, perhaps a facet-wrapped histogram of size distributions?
# Probably should make it look prettier. Let's eventually pick a color theme on color brewer and coordinate all of our visuals! 
size_hist <- ggplot(lsize_expanded, aes(x = SIZE)) +
  geom_histogram(bins = 15, aes(fill = SITE)) +
  facet_wrap(~ SITE, scale = "free") + 
  theme_classic() +
  theme(legend.position = "none") + 
  scale_y_continuous(expand = c(0,0))

size_hist
# Looks normal, how about a qq?

size_qq <- ggplot(lsize_expanded, aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~ SITE, scale = "free")

# Looks pretty normal

# Use t-tests to compare means

# Starting with MPA - IVEE

# Creating Vectors for F and T-test

IVEE_2012 <- lsizempa_expanded %>% 
  filter(SITE == "IVEE", YEAR == "2012") %>% 
  pull(SIZE)

IVEE_2017 <- lsizempa_expanded %>% 
  filter(YEAR == "2017",SITE == "IVEE") %>% 
  pull(SIZE)

# F Test
# H0: Variances are EQUAL
# HA: Variances are NOT equal

IVEE_ftest <- var.test(IVEE_2012, IVEE_2017)
IVEE_ftest

# P value = .307, so variances are considered EQUAL

# T-Test for IVEE
# H0: Samples are drawn from populations with the same mean
# HA: Samples are drawn from populations with different means

t_IVEE <- t.test(IVEE_2012, IVEE_2017, var.equal = TRUE)
t_IVEE

# P-value = .0599
# Retain Null that the samples are drawn from populations with the same means
# Sample Means (2012 = 66.07, 2017 = 71.45)

# Effect Size

eff_size_IVEE <- cohen.d(IVEE_2012, IVEE_2017)
eff_size_IVEE

# d-estimate: -0.377 (small)

# power test

power_IVEE <- pwr.t.test(n = 632, d = .377, sig.level = 0.05, power = NULL)
power_IVEE

# power = .999



# MPA - NAPL

# Creating Vectors for F and T-Tests

NAPL_2012 <- lsizempa_expanded %>% 
  filter(YEAR == "2012", SITE == "NAPL") %>% 
  pull(SIZE)

NAPL_2017 <- lsizempa_expanded %>% 
  filter(YEAR == "2017", SITE == "NAPL") %>% 
  pull(SIZE)

# F-Test
# H0: Variances are EQUAL
# HA: Variances are NOT equal

NAPL_ftest <- var.test(NAPL_2012, NAPL_2017)
NAPL_ftest

# P-value = 0.7685, so variances are considered EQUAL

# T-Test for NAPL
# H0: Samples are drawn from populations with the same mean
# HA: Samples are drawn from populations with different means

t_NAPL <- t.test(NAPL_2012, NAPL_2017, var.equal = TRUE)
t_NAPL

# P-Value = .5002
# Retain Null, samples are drawn from populations with the same means
# Sample means (2012 = 73, 2017 = 76.23)

# Effect Size

eff_size_NAPL <- cohen.d(NAPL_2012,NAPL_2017)
eff_size_NAPL

# d-estimate: -0.283 (small)

# power test

power_NAPL <- pwr.t.test(n = 118, d = .283, sig.level = 0.05, power = NULL)
power_NAPL

# power = .581



# NON-MPA

# NON MPA - AQUE

# Creating Vectors for F and T-Tests

AQUE_2012 <- lsizempa_expanded %>% 
  filter(YEAR == "2012", SITE == "AQUE") %>% 
  pull(SIZE)

AQUE_2017 <- lsizempa_expanded %>% 
  filter(YEAR == "2017", SITE == "AQUE") %>% 
  pull(SIZE)

# F-Test
# H0: Variances are EQUAL
# HA: Variances are NOT equal

AQUE_ftest <- var.test(AQUE_2012, AQUE_2017)
AQUE_ftest

# P-value = 0.298, so variances are considered EQUAL

# T-Test for NAPL
# H0: Samples are drawn from populations with the same mean
# HA: Samples are drawn from populations with different means

t_AQUE <- t.test(AQUE_2012, AQUE_2017, var.equal = TRUE)
t_AQUE

# P-Value = .2097
# Retain Null, samples are drawn from populations with the same means
# Sample means (2012 = 71, 2017 = 73.89)

# Effect Size

eff_size_AQUE <- cohen.d(AQUE_2012, AQUE_2017)
eff_size_AQUE

# d-estimate: -0.256 (small)

# power test

power_AQUE <- pwr.t.test(n = 105, d = .256, sig.level = 0.05, power = NULL)
power_AQUE

# power = .4548



# NON MPA - CARP

# Creating Vectors for F and T-Tests

CARP_2012 <- lsizempa_expanded %>% 
  filter(YEAR == "2012", SITE == "CARP") %>% 
  pull(SIZE)

CARP_2017 <- lsizempa_expanded %>% 
  filter(YEAR == "2017", SITE == "CARP") %>% 
  pull(SIZE)

# F-Test
# H0: Variances are EQUAL
# HA: Variances are NOT equal

CARP_ftest <- var.test(CARP_2012, CARP_2017)
CARP_ftest

# P-value = 0.204, so variances are considered EQUAL

# T-Test for CARP
# H0: Samples are drawn from populations with the same mean
# HA: Samples are drawn from populations with different means

t_CARP <- t.test(CARP_2012, CARP_2017, var.equal = TRUE)
t_CARP

# P-Value = .1819
# Retain Null, samples are drawn from populations with the same means
# Sample means (2012 = 74.36, 2017 = 72.23)

# Effect Size

eff_size_CARP <- cohen.d(CARP_2012, CARP_2017)
eff_size_CARP

# d-estimate: 0.159 (negligible)

# power test

power_CARP <- pwr.t.test(n = 783, d = .159, sig.level = 0.05, power = NULL)
power_CARP

# power = .882



# NON MPA - MOHK

# Creating Vectors for F and T-Tests

MOHK_2012 <- lsizempa_expanded %>% 
  filter(YEAR == "2012", SITE == "MOHK") %>% 
  pull(SIZE)

MOHK_2017 <- lsizempa_expanded %>% 
  filter(YEAR == "2017", SITE == "MOHK") %>% 
  pull(SIZE)

# F-Test
# H0: Variances are EQUAL
# HA: Variances are NOT equal

MOHK_ftest <- var.test(MOHK_2012, MOHK_2017)
MOHK_ftest

# P-value = 0.1509, so variances are considered EQUAL

# T-Test for NAPL
# H0: Samples are drawn from populations with the same mean
# HA: Samples are drawn from populations with different means

t_MOHK <- t.test(MOHK_2012, MOHK_2017, var.equal = TRUE)
t_MOHK

# P-Value = 6.27 e-05
# Reject Null, samples are drawn from populations with different means
# Sample means (2012 = 77.25, 2017 = 72)

# One-Sided T-Test for NAPL
# H0: The mean lobster size in 2012 is not greater than mean lobster size in 2017
# HA: The mean lobster size in 2012 is greater than the mean lobster size in 2017

onesided_t_MOHK <- t.test(MOHK_2012, MOHK_2017, var.equal = TRUE, alternative = "greater")
onesided_t_MOHK

# P-Value = 3.138e-05
# Reject Null, mean lobster size in 2012 is greater than mean lobster size in 2017

# Effect Size

eff_size_MOHK <- cohen.d(MOHK_2012, MOHK_2017)
eff_size_MOHK

# d-estimate: 0.5408 (medium)

# power test

power_MOHK <- pwr.t.test(n = 261, d = .5408, sig.level = 0.05, power = NULL)
power_MOHK

# power = .999


# Summary Table of means at location, each year

summary_lobster_size <- lsizempa_expanded %>% 
  group_by(SITE, YEAR) %>% 
  summarize(mean_lobs_size = mean(SIZE), sd = sd(SIZE), sample_size = length(SIZE))

summary_lobster_size

sumlobs <- read_csv("Mean Lobster Size Table.csv")

lobs_size_table <- kable(sumlobs, col.names = c(" ", "2012", "2017", " "), align = 'c') %>% 
  kable_styling(font_size = 11, "condensed") %>% 
  group_rows("MPA", 1, 2) %>% 
  group_rows("Non-MPA", 3, 5) %>% 
  add_header_above(c(" " = 1, "Mean ± Standard Deviation" = 2, "Difference" = 1), italic = T, align = 'c') %>% 
  add_header_above(c(" " = 1, "Measured Lobster Carapace Length (mm)" = 3), bold = T, align = 'c') %>% 
  column_spec(1, width = "1cm") %>% 
  column_spec(2, width = "2cm") %>% 
  column_spec(3, width = "2cm") %>% 
  column_spec(4, width = "1cm")
  


lobs_size_table
```


*Legal and Illegal Lobster Trapping in 2017*
```{r, echo=FALSE}
#chi-square test for independance
#The legal minimum carapace size for lobster is 82.6 mm. What proportion of observed lobsters at each site are above the legal minimum? Does that proportion differ significantly across the 5 sites? 
#first, set up a table of proportions for 2017. The table should have the sites as the the row names, and then the columns shoulf be the proportions below and above the legal minimum carapace size (82.6mm)

lsize_prop_table <- lsize_expanded %>% 
  select(-YEAR, -COUNT) %>% 
  mutate(Legality = ifelse(SIZE > 82.6, "Above Legal Minimum", "Below Legal Minimum")) %>% 
  count(SITE, Legality) %>% 
  spread(Legality, n) %>% 
  select(-SITE)

rownames(lsize_prop_table) <- c("AQUE", "CARP", "IVEE", "MOHK", "NAPL")

lsize_prop_table

#H0:There is no different in proportions
#HA: There is a difference in proportions

#look at the actual proportions

lob_prop <- prop.table(as.matrix(lsize_prop_table), 1)

#run a chi-square test for independence 

lsize_x2 <- chisq.test(lsize_prop_table)
lsize_x2

#p-value = 0.0009864, reject the null! We have significant evidence to suggest there is a difference in proportions of lobsters above/below the legal size limit between sites
```

 #in console, running lsize_x2$stdres to see which sites differ significantly:
 Above Legal Minimum Below Legal Minimum
AQUE           0.1464223          -0.1464223
CARP           1.8631463          -1.8631463
IVEE          -1.2357993           1.2357993
MOHK          -3.2327773           3.2327773
NAPL           2.5706474          -2.5706474

Standardized residuals greater than |2| indicate significance (I think?)

###So, we can say that the proportions between legal/illegal lobster size is differant at MOHK and NAPL. At MOHK, 87% of lobsters observed were smaller than the legal limit. At NAPL (an MPA), only 66% of lobsters were below the legal limit. The other three sites had between 74-78% of their lobsters below the legal size limit. 

```{r, echo=FALSE}
#Making a neat Kable Table of the proportions 


lob_prob_table <- kable(lob_prop, digits = 2) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) 

lob_prob_table

```



```{r}
#map <- read_csv("Long Lat.csv")
lobster_icon <- makeIcon("small_lob.png", "small_lob@2x.png", 18, 40, background = FALSE)
my_map <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(icon = lob_icon, lng=map$Longtitude, lat=map$Latitude, popup="LTER Site") 
  
lob_icon <- makeIcon("http://www.stickpng.com/assets/thumbs/585d3b7ecb11b227491c3355.png", 18, 40)

my_map
```

##CONCLUSION
#####
The California Spiny Lobster, a keystone species through the california kelp forest ecosystems, also supports an economically important commercial and recreational fishery. In order to preserve lobster populations, California established multiple marine protected areas (MPAs) in 2012 to create “safe zones” for lobsters to spawn undisturbed by fishing pressure. This report aims to analyze the differences in lobster size, abundance, and fishing pressure by site, specifically comparing the differences between MPA sites and non-MPA sites. The following statements summarize the findings of this report:

Statement 1
Statement 2
Statement 3
Statement 4

In order to stay up to date on management best practices, this report should be updated yearly with new data, especially since many variables other than fishing pressure can have a large and immediate impact on loster populations, such as anomalies in water temperature from the Pacific Decadal Oscillation. Fishery managers should know as soon as possible when there are major disturbances to lobster populations in order to make well-informed decisions about how to optimize the lobster fishery for both economic value and conservation. Given that the establishment of the southern California MPAs are fairly recent, longer term monitoring is needed to determine if they boost net lobster populations, or create detrimental pressure to non-MPA zones. Studies such as the SBC LTER will be essential in monitoring these changes going forward. 



##REFERENCES

#####
1. California Department of Fish and Wildlife, Marine Region. “California Spiny Lobster Fishery Management Plan.” California Spiny Lobster Fishery Management Plan, 13 Apr. 2016. www.wildlife.ca.gov/Conservation/Marine/Lobster-FMP.

2. California Department of Fish and Wildlife Marine Region: Invertebrate Project. “Spiny Lobster Fishery Management Plan Harvest Control Rule.” Spiny Lobster Fishery Management Plan Harvest Control Rule, 9 Apr. 2018. nrm.dfg.ca.gov/FileHandler.ashx?DocumentID=156078&inline.

3. Guenther, Carla M., et al. “Trophic Cascades Induced by Lobster Fishing Are Not Ubiquitous in Southern California Kelp Forests.” PLoS ONE, vol. 7, no. 11, 2012, doi:10.1371/journal.pone.0049396.

4. Reed, D. . 2017. SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (Panulirus interruptus), ongoing since 2012. Santa Barbara Coastal Long 

